Exploratory Data Analysis

### 2023-11-20 analysis/aggressive_care.qmd
who knows

## Exploratory Data Analysis

Loading the data from ATACH-2 and ERICH, and filtering out the variables needed for the analysis.

```{r loading data}

# load data
data <- read_rds(here("data", "all.rds"), refhook = NULL)

# filter out only ERICH and ATACH2 and keep relevant variables
cleaned_data <- data |>
  mutate(id = 1:n()) |>
  filter(study == "ERICH" | study == "ATACH-2") |>
  select(id, study, ich_laterality, age, sex, race, ethnicity, sbp_baseline, gcs_baseline, ich_volume_baseline, ivh, ich_location, neurosurgery_evac, evd, tracheostomy, peg, comfort_care_binary, comfort_care_day) |>
  drop_na(ich_laterality, neurosurgery_evac)

Now we can summarize the preliminary data.

```{r exploratory data analysis}

counts <- cleaned_data |>
  tabyl(ich_laterality, neurosurgery_evac, ich_location) |>
  adorn_percentages("row") |>
  adorn_pct_formatting(rounding = "half up", digits = 1)

counts

counts_tidy <- cleaned_data |>
  group_by(ich_laterality, neurosurgery_evac) |>
  summarize(n = n()) |>
  mutate(
    total = sum(n),
    prop = n / sum(n)
  ) |>
  ungroup()

counts_tidy

ggplot(cleaned_data) +
  geom_mosaic(
    aes(
      x = product(neurosurgery_evac, ich_laterality),
      fill = ich_laterality, alpha = neurosurgery_evac
    ),
    divider = mosaic("v")
  )

@@

>>>>

@@

```{r neurosurgery, warning=FALSE, message=FALSE}

table_two_vars <- c("neurosurgery_evac", "evd", "days_mechanical_ventilation", "tracheostomy", "ich_laterality", "comfort_care_binary")

table2 <- data_erich_atach |>
  select(table_two_vars) |>
  tbl_summary(
    by = ich_laterality,
    missing = "no",
    label = list(
      neurosurgery_evac ~ "Neurosurgical Intervention",
      evd ~ "Ventriculostomy",
      days_mechanical_ventilation ~ "Intubation Days",
      tracheostomy ~ "Tracheostomy",
      comfort_care_binary ~ "Comfort Care"
    )
  ) |>
  add_p() |>
  add_overall() |>
  bold_labels()

table2

## Neurosurgical Intervention - Multivariate Regression

```{r neurosurgical intervention, warning=FALSE, message=FALSE}
#| fig-width: 10

neurosurgery_glm <- glm(
  neurosurgery_evac ~ ich_laterality + age + ich_location + gcs_baseline + ich_volume_baseline + ivh + time_symptoms_to_ed,
  data = data_erich_atach,
  family = binomial
)

neurosurgery_glm |>
  tbl_regression(
    exponentiate = TRUE,
    include = ich_laterality
  )

vent_glm <- glm(
  days_mechanical_ventilation ~ ich_laterality + age + ich_location + gcs_baseline + ich_volume_baseline + ivh + time_symptoms_to_ed,
  data = data_erich_atach,
  family = poisson
)

vent_glm |>
  tbl_regression(
    exponentiate = TRUE,
    include = ich_laterality
  )

comfort_glm <- glm(
  comfort_care_binary ~ ich_laterality + age + ich_location + gcs_baseline + ich_volume_baseline + ivh + time_symptoms_to_ed,
  data = data_erich_atach,
  family = binomial
)

comfort_glm |>
  tbl_regression(
    exponentiate = TRUE,
    include = ich_laterality
  )

### 2023-11-20 analysis/eda.qmd
who knows

- Collapse some categorical values
- IVH as binary instead of volume

@@

```{r table1, warning=FALSE, message=FALSE}
#| fig-width: 10

table_one_vars <- c("age", "sex", "race", "ethnicity", "time_symptoms_to_ed", "sbp_baseline", "nihss_baseline", "gcs_baseline", "ich_laterality", "htn", "dm2", "stroke", "tobacco", "ich_location", "ich_volume_baseline", "ivh", "study")

@@

```{r neurosurgery, warning=FALSE, message=FALSE}

table_two_vars <- c("neurosurgery_evac", "evd", "days_mechanical_ventilation", "tracheostomy", "ich_laterality", "comfort_care_binary")

table2 <- data_erich_atach |>
  select(table_two_vars) |>
  tbl_summary(
    by = ich_laterality,
    missing = "no",
    label = list(
      neurosurgery_evac ~ "Neurosurgical Intervention",
      evd ~ "Ventriculostomy",
      days_mechanical_ventilation ~ "Intubation Days",
      tracheostomy ~ "Tracheostomy",
      comfort_care_binary ~ "Comfort Care"
    )
  ) |>
  add_p() |>
  add_overall() |>
  bold_labels()

table2

## Neurosurgical Intervention - Multivariate Regression

```{r neurosurgical intervention, warning=FALSE, message=FALSE}
#| fig-width: 10

neurosurgery_glm <- glm(
  neurosurgery_evac ~ ich_laterality + age + ich_location + gcs_baseline + ich_volume_baseline + ivh + time_symptoms_to_ed,
  data = data_erich_atach,
  family = binomial
)

neurosurgery_glm |>
  tbl_regression(
    exponentiate = TRUE,
    include = ich_laterality
  )

left_right_regression <- function(name, x, family) {
  name <- glm(
    {{ x }} ~ ich_laterality + age + ich_location + gcs_baseline + ich_volume_baseline + ivh + time_symptoms_to_ed,
    data = data_erich_atach,
    family = family
  )

  {{ x }} |>
    tbl_regression(
      exponentiate = TRUE,
      include = ich_laterality
    )
}

left_right_regression(neurosurgery_glm, neurosurgery_evac, binomial)

vent_glm <- glm(
  days_mechanical_ventilation ~ ich_laterality + age + ich_location + gcs_baseline + ich_volume_baseline + ivh + time_symptoms_to_ed,
  data = data_erich_atach,
  family = poisson
)

model_parameters(vent_glm,
  centrality = "mean", dispersion = TRUE,
  test = FALSE, verbose = FALSE, ci_method = "hdi", priors = TRUE,
  exponentiate = TRUE
) |>
  print(digits = 2, zap_small = TRUE)

## Outcomes

```{r euroqol, warning=FALSE, message=FALSE}
#| fig-width: 10

data <- read_csv(here("data", "all.csv")) |>
  filter(ich_location == c("Lobar", "Basal Ganglia", "Thalamus")) |>
  filter(!is.na(ich_laterality))

data |>
  ggplot(aes(euro_vas_90, fill = ich_laterality)) +
  geom_density(color = NA, alpha = .8)

>>>>

- Collapse some categorical values
- IVH as binary instead of volume
- Some percentages values are off, likely due to missing data (e.g., Hypertension)

@@

```{r table1, warning=FALSE, message=FALSE}
#| fig-width: 10
#|
table_one_vars <- c("age", "sex", "race", "ethnicity", "time_symptoms_to_ed", "sbp_baseline", "nihss_baseline", "gcs_baseline", "ich_laterality", "htn", "dm2", "stroke", "tobacco", "ich_location", "ich_volume_baseline", "ivh", "study")

@@
